---
title: "A Brief Introducton to Archaeological Data Carpentry"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

We are going to do three things:

1. Learn the basics of R Markdown and knitting a document
2. Learn the basics of five main verbs of data manipulation
3. Learn the basics of four common types of data visualisation 

We will use packages in the tidyverse, a set of R packages for data science. All packages share an underlying design philosophy, grammar, and data structures: https://www.tidyverse.org/

If you want to browse more packages that are useful for archaeology, we have a list here: https://github.com/benmarwick/ctv-archaeology

# Getting the data

We will download some real archaeological data that has been published in this journal article: 

> Marwick, Ben, et al. "Early modern human lithic technology from Jerimalai, East Timor." _Journal of Human Evolution_ 101 (2016): 45-64. https://doi.org/10.1016/j.jhevol.2016.09.004

The data have been deposites on Figshare.com, a major multi-disciplinary data repository (free to use).

In this chunk below we will:

1. download the data in a zip file
2. unzip this zip file to access the individual files
3. delete the zip file
 
```{r cache=TRUE}
# get data from the public repository
download.file("https://ndownloader.figshare.com/files/6042231",
              destfile = "data/Marwick-et-al_2016-JHE.zip")

# unzip the downloaded zip file so we can access the individual files 
unzip("data/Marwick-et-al_2016-JHE.zip",
      exdir = "data")

# delete the zip file
unlink("data/Marwick-et-al_2016-JHE.zip")
```

Now the data have been downloaded onto our hard drive, we can load one spreadsheet into our current R session to begin working on it:

```{r message=FALSE, warning=FALSE}
# load the packages of the tidyverse
library(tidyverse)

# load the data
jerimalai_lithics_square_a <- 
  read_csv("data/Pleistocene-aged-stone-artefacts-from-Jerimalai--East-Timor-master/manuscript/data/Jerimalai_All_Artefacts_Square_A.csv")
```

# Five Main Verbs for Data Manipulation

We are use the dplyr package from the tidyverse. It has an awesome cheat-sheet here: https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

## Filter

```{r}

```

## Mutate

```{r}

```

## Arrange

```{r}

```

## Summarise

```{r}

```

## Group by

```{r}

```

# Four Common Types of Data Visualisation 

Here we are using the ggplot2 package, here's the cheatsheet: https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf

## Histogram

```{r}

```

## Bar plot

```{r}

```

## Boxplot

```{r}

```

## Scatterplot

```{r}

```

How to save a plot as an image file?



